---
title: "Shiny App Development"
author: "Kyle Knox"
date: "Week 12"
output:
  html_document: 
    toc: no
    toc_depth: 4
    toc_float: yes
    fig_width: 6
    number_sections: yes
    toc_collapsed: yes
    code_folding: hide
    code_download: yes
    smooth_scroll: true
    theme: readable
    fig_height: 8
    runtime: shiny
---
  
  <style type="text/css">
  
  div#TOC li {
list-style:none;
background-color:lightgray;
background-image:none;
background-repeat:none;
background-position:0;
font-family: Arial, Helvetica, sans-serif;
color: #780c0c;
  }

/* mouse over link */
  div#TOC a:hover {
color: red;
}

/* unvisited link */
  div#TOC a:link {
color: blue;
}



h1.title {
  font-size: 24px;
  color: Darkblue;
  text-align: center;
  font-family: Arial, Helvetica, sans-serif;
  font-variant-caps: normal;
}
h4.author { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 {
  font-size: 22px;
  font-family: "Times New Roman", Times, serif;
  color: darkred;
  text-align: center;
}
h2 { /
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: navy;
  text-align: left;
}

h3 { 
  font-size: 15px;
  font-family: "Times New Roman", Times, serif;
  color: navy;
  text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: darkred;
  text-align: left;
}

/* unvisited link */
  a:link {
    color: green;
  }

/* visited link */
  a:visited {
    color: green;
  }

/* mouse over link */
  a:hover {
    color: red;
  }

/* selected link */
  a:active {
    color: yellow;
  }

</style>
  
  
```{r setup, include=FALSE}
# code chunk specifies whether the R code, warnings, and output 
# will be included in the output files.
options(repos = list(CRAN="http://cran.rstudio.com/"))

if (!require("tidyverse")) {
  install.packages("tidyverse")
  library(tidyverse)
}
if (!require("knitr")) {
  install.packages("knitr")
  library(knitr)
}
if (!require("cowplot")) {
  install.packages("cowplot")
  library(cowplot)
}
if (!require("latex2exp")) {
  install.packages("latex2exp")
  library(latex2exp)
}
if (!require("plotly")) {
  install.packages("plotly")
  library(plotly)
}
if (!require("gapminder")) {
  install.packages("gapminder")
  library(gapminder)
}
if (!require("png")) {
  install.packages("png")             # Install png package
  library("png")
}
if (!require("RCurl")) {
  install.packages("RCurl")             # Install RCurl package
  library("RCurl")
}
if (!require("colourpicker")) {
  install.packages("colourpicker")              
  library("colourpicker")
}
if (!require("gifski")) {
  install.packages("gifski")              
  library("gifski")
}
if (!require("magick")) {
  install.packages("magick")              
  library("magick")
}
if (!require("grDevices")) {
  install.packages("grDevices")              
  library("grDevices")
}
### ggplot and extensions
if (!require("ggplot2")) {
  install.packages("ggplot2")              
  library("ggplot2")
}
if (!require("gganimate")) {
  install.packages("gganimate")              
  library("gganimate")
}
if (!require("ggridges")) {
  install.packages("ggridges")              
  library("ggridges")
}
if (!require("dplyr")) {
  install.packages("dplyr")              
  library("dplyr")
}
if (!require("htmltools")) {
  install.packages("htmltools")              
  library("htmltools")
}
if (!require("rmarkdown")) {
  install.packages("rmarkdown")              
  library("rmarkdown")
}

if (!require("sf")) {
  install.packages("sf")
  library("sf")
}
if (!require("geojsonio")) {
  install.packages("geojsonio")
  library("geojsonio")
}
if (!require("leaflet")) {
  install.packages("leaflet")
  library("leaflet")
}
if (!require("shiny")) {
  install.packages("shiny")
  library("shiny")
}
# knitr::opts_knit$set(root.dir = "/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA533/week10)

knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE)
```

```{r, fig.height=15, fig.width=12}
library(shiny)

ui <- fluidPage(
  withMathJax(),
  tags$div(HTML("<script type='text/x-mathjax-config'>
            MathJax.Hub.Config({
            tex2jax: {inlineMath: [['$','$'], ['\\(','\\)']]}
            });
            </script>")),
  br(), br(),
  wellPanel(style = "background-color:#acdfe6;",
            titlePanel(h2("Exponential Distribution of Random Variables", align = "center",
                          style = "color:darkred; font-family:verdana; 
                      font-variant: small-caps; font-weight: bold;"))
  ),
  sidebarPanel(style = "background-color:#acdfe6;",
               tags$div( h4("User Input Panel", align = "center",
                          style = "color:darkred; font-family:verdana; 
                      font-variant: small-caps; font-weight: bold;")),
               br(),
               sliderInput("obs",
                           "Number of observations (n)",
                           min = 1,
                           max = 1000,
                           value = 100,
                           animate = list(interval=550, loop=T)),
               sliderInput("rate",
                           "Rate parameter (lambda)",
                           min = 0.1,
                           max = 5,
                           value = 1,
                           step = 0.1),
               br()
  ),
  mainPanel(
    style = "background-color:lightgray; text-align: center;",
    plotOutput("histPlot", height = "300px"),
    plotOutput("densityPlot", height = "300px")
  )
)

server <- function(input, output) {
  output$histPlot <- renderPlot({
    set.seed(3)
    dist <- rexp(input$obs, rate = input$rate)
    hist(dist,
         col = "darkgreen",
         main = "Histogram of Exponentially Distributed Random Variables",
         xlab = "Values",
         ylab = "Frequency")
  })

  output$densityPlot <- renderPlot({
    set.seed(3)
    dist <- rexp(input$obs, rate = input$rate)
    plot(density(dist),
         main = "Density Plot of Exponentially Distributed Random Variables",
         xlab = "Values",
         ylab = "Density",
         col = "red")
    lines(density(dist), col = "red", lwd = 2)
  })
}

shinyApp(ui = ui, server = server)

```
