---
title: "Life Expectancy and Income for the Year 2015"
author: "Kyle Knox"
date: "Week 6"
output:
  html_document: 
  toc: yes
  toc_depth: 4
  toc_float: yes
  fig_width: 6
  number_sections: yes
  toc_collapsed: yes
  code_folding: hide
  code_download: yes
  smooth_scroll: true
  theme: readable
  fig_height: 4
---
  
  <style type="text/css">
  
  div#TOC li {
list-style:none;
background-color:lightgray;
background-image:none;
background-repeat:none;
background-position:0;
font-family: Arial, Helvetica, sans-serif;
color: #780c0c;
  }

/* mouse over link */
  div#TOC a:hover {
color: red;
}

/* unvisited link */
  div#TOC a:link {
color: blue;
}



h1.title {
  font-size: 24px;
  color: Darkblue;
  text-align: center;
  font-family: Arial, Helvetica, sans-serif;
  font-variant-caps: normal;
}
h4.author { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 {
  font-size: 22px;
  font-family: "Times New Roman", Times, serif;
  color: darkred;
  text-align: center;
}
h2 { /
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: navy;
  text-align: left;
}

h3 { 
  font-size: 15px;
  font-family: "Times New Roman", Times, serif;
  color: navy;
  text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: darkred;
  text-align: left;
}

/* unvisited link */
  a:link {
    color: green;
  }

/* visited link */
  a:visited {
    color: green;
  }

/* mouse over link */
  a:hover {
    color: red;
  }

/* selected link */
  a:active {
    color: yellow;
  }

</style>
  
  
  ```{r setup, include=FALSE}
# code chunk specifies whether the R code, warnings, and output 
# will be included in the output files.
options(repos = list(CRAN="http://cran.rstudio.com/"))
if (!require("tidyverse")) {
  install.packages("tidyverse")
  library(tidyverse)
}
if (!require("knitr")) {
  install.packages("knitr")
  library(knitr)
}
if (!require("cowplot")) {
  install.packages("cowplot")
  library(cowplot)
}
if (!require("latex2exp")) {
  install.packages("latex2exp")
  library(latex2exp)
}
if (!require("plotly")) {
  install.packages("plotly")
  library(plotly)
}
if (!require("gapminder")) {
  install.packages("gapminder")
  library(gapminder)
}
if (!require("png")) {
  install.packages("png")             # Install png package
  library("png")
}
if (!require("RCurl")) {
  install.packages("RCurl")             # Install RCurl package
  library("RCurl")
}
if (!require("colourpicker")) {
  install.packages("colourpicker")              
  library("colourpicker")
}
if (!require("gifski")) {
  install.packages("gifski")              
  library("gifski")
}
if (!require("magick")) {
  install.packages("magick")              
  library("magick")
}
if (!require("grDevices")) {
  install.packages("grDevices")              
  library("grDevices")
}
### ggplot and extensions
if (!require("ggplot2")) {
  install.packages("ggplot2")              
  library("ggplot2")
}
if (!require("gganimate")) {
  install.packages("gganimate")              
  library("gganimate")
}
if (!require("ggridges")) {
  install.packages("ggridges")              
  library("ggridges")
}
# knitr::opts_knit$set(root.dir = "/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA533/week6)
# knitr::opts_knit$set(root.dir = "C:\\STA490\\w05")

knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE)
```
```{r code, fig.width=8, fig.height=4.8, echo=FALSE}
income <- read.csv("/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA533/week5/income_per_person.csv", check.names = FALSE)

incomepp <- income %>%
  gather(key = "year", value = "income", -geo) %>%
  rename(country = geo)

print(incomepp)

life <- read.csv("/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA533/week5/life_expectancy_years.csv", check.names = FALSE)

lifeyr <- life %>%
  gather(key = "year", value = "income", -geo) %>%
  rename(country = geo)

print(lifeyr)

LifeExpIncom <- inner_join(incomepp, lifeyr, by = c("country", "year"))

LifeExpIncom <- LifeExpIncom %>%
  rename(income = income.x, lifeExp = income.y)

print(LifeExpIncom)

countryR <- read.csv("/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA533/week5/countries_total.csv", check.names = FALSE)
PopT <- read.csv("/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA533/week5/population_total.csv", check.names = FALSE)

country_year <- countryR %>%
  rename(country = name)
country_region <- country_year %>%
  select(country, region)

LifeIncomeRegion <- inner_join(LifeExpIncom, country_region, by = "country")

population_fixed <- PopT %>%
  gather(key = "year", value = "population", -geo) %>%
  rename(country = geo) %>%
  mutate(year = as.numeric(year))

LifeIncomeRegion$year <- as.numeric(LifeIncomeRegion$year)

LifeIncomePopulation <- inner_join(LifeIncomeRegion, population_fixed, by = c("country", "year"))

write.csv(LifeIncomePopulation, "/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA533/week5/LifeIncomePopulation.csv", row.names = FALSE)

data2000 <- LifeIncomePopulation %>%
  filter(year == 2015)

write.csv(LifeIncomePopulation, "/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA533/week5/data2000.csv", row.names = FALSE)


```
