---
title: "Reference Map: Philadelphia Shootings"
author: "Kyle Knox"
date: "Week 9"
output:
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
    fig_width: 6
    number_sections: yes
    toc_collapsed: yes
    code_folding: hide
    code_download: yes
    smooth_scroll: true
    theme: readable
    fig_height: 4
---
  
  <style type="text/css">
  
  div#TOC li {
list-style:none;
background-color:lightgray;
background-image:none;
background-repeat:none;
background-position:0;
font-family: Arial, Helvetica, sans-serif;
color: #780c0c;
  }

/* mouse over link */
  div#TOC a:hover {
color: red;
}

/* unvisited link */
  div#TOC a:link {
color: blue;
}



h1.title {
  font-size: 24px;
  color: Darkblue;
  text-align: center;
  font-family: Arial, Helvetica, sans-serif;
  font-variant-caps: normal;
}
h4.author { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 {
  font-size: 22px;
  font-family: "Times New Roman", Times, serif;
  color: darkred;
  text-align: center;
}
h2 { /
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: navy;
  text-align: left;
}

h3 { 
  font-size: 15px;
  font-family: "Times New Roman", Times, serif;
  color: navy;
  text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: darkred;
  text-align: left;
}

/* unvisited link */
  a:link {
    color: green;
  }

/* visited link */
  a:visited {
    color: green;
  }

/* mouse over link */
  a:hover {
    color: red;
  }

/* selected link */
  a:active {
    color: yellow;
  }

</style>
  
  
```{r setup, include=FALSE}
# code chunk specifies whether the R code, warnings, and output 
# will be included in the output files.
options(repos = list(CRAN="http://cran.rstudio.com/"))

if (!require("tidyverse")) {
  install.packages("tidyverse")
  library(tidyverse)
}
if (!require("knitr")) {
  install.packages("knitr")
  library(knitr)
}
if (!require("cowplot")) {
  install.packages("cowplot")
  library(cowplot)
}
if (!require("latex2exp")) {
  install.packages("latex2exp")
  library(latex2exp)
}
if (!require("plotly")) {
  install.packages("plotly")
  library(plotly)
}
if (!require("gapminder")) {
  install.packages("gapminder")
  library(gapminder)
}
if (!require("png")) {
  install.packages("png")             # Install png package
  library("png")
}
if (!require("RCurl")) {
  install.packages("RCurl")             # Install RCurl package
  library("RCurl")
}
if (!require("colourpicker")) {
  install.packages("colourpicker")              
  library("colourpicker")
}
if (!require("gifski")) {
  install.packages("gifski")              
  library("gifski")
}
if (!require("magick")) {
  install.packages("magick")              
  library("magick")
}
if (!require("grDevices")) {
  install.packages("grDevices")              
  library("grDevices")
}
### ggplot and extensions
if (!require("ggplot2")) {
  install.packages("ggplot2")              
  library("ggplot2")
}
if (!require("gganimate")) {
  install.packages("gganimate")              
  library("gganimate")
}
if (!require("ggridges")) {
  install.packages("ggridges")              
  library("ggridges")
}
if (!require("dplyr")) {
  install.packages("dplyr")              
  library("dplyr")
}
if (!require("htmltools")) {
  install.packages("htmltools")              
  library("htmltools")
}
if (!require("rmarkdown")) {
  install.packages("rmarkdown")              
  library("rmarkdown")
}

if (!require("sf")) {
  install.packages("sf")
  library("sf")
}
if (!require("geojsonio")) {
  install.packages("geojsonio")
  library("geojsonio")
}
if (!require("leaflet")) {
  install.packages("leaflet")
  library("leaflet")
}
# knitr::opts_knit$set(root.dir = "/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA533/week9)

knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE)
```

```{r code, fig.height= 7, fig.width= 9}
shootings <- st_read("https://raw.githubusercontent.com/kyleknox3/STA553/main/week9/shootings.geojson");

shootings_2023 <- shootings %>%
  filter(year == 2023)

shootings_2023 <- shootings_2023 %>%
  mutate(fatal = ifelse(fatal == 1, "Yes", "No"))

map_title <- "PHL Shootings 2023"

main_map <- leaflet(data = shootings_2023) %>%
  addProviderTiles(providers$CartoDB.DarkMatter, group="Dark") %>%
  addProviderTiles(providers$CartoDB.DarkMatterNoLabels, group="DarkLabel") %>%  
  addProviderTiles(providers$Esri.NatGeoWorldMap, group="Esri") %>%
  #addTiles(group = "Dark") %>%
  #addTiles(group = "DarkLabel", options = providerTileOptions(noWrap = TRUE)) %>%
  #addTiles(group = "Esri") %>%
  addCircleMarkers(
    radius = ~ifelse(fatal == "Yes", 5, 3),
    color = ~case_when(
      wound == "Head" ~ "red",
      wound == "Multiple" ~ "blue",
      wound == "Leg" ~ "green",
      wound == "Neck" ~ "purple",
      TRUE ~ "yellow"
    ),
    popup = ~paste("Fatal: ", fatal, 
                   "<br>Wound: ", wound, 
                   "<br>Age: ", age, 
                   "<br>Race: ", race, 
                   "<br>Location: ", location),
    fillOpacity = 0.8
  ) %>%
  addLayersControl(baseGroups = c('Dark', 'DarkLabel', 'Esri'),
                   options = layersControlOptions(collapsed = TRUE)) %>%
  addLegend(
    position = "bottomleft",
    colors = c("red", "blue", "green", "purple", "yellow"),
    labels = c("Head", "Multiple", "Leg", "Neck", "Other"),
    title = "Legend"
  ) %>%
  addControl(
    html = sprintf('<div class="leaflet-control-attribution leaflet-control" style="font-size: 20px;">%s</div>', map_title),
    position = "topleft"
  )

main_map
```