---
title: "Reference Map: Philadelphia Shootings"
author: "Kyle Knox"
date: "Week 9"
output:
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
    fig_width: 6
    number_sections: yes
    toc_collapsed: yes
    code_folding: hide
    code_download: yes
    smooth_scroll: true
    theme: readable
    fig_height: 4
---
  
  <style type="text/css">
  
  div#TOC li {
list-style:none;
background-color:lightgray;
background-image:none;
background-repeat:none;
background-position:0;
font-family: Arial, Helvetica, sans-serif;
color: #780c0c;
  }

/* mouse over link */
  div#TOC a:hover {
color: red;
}

/* unvisited link */
  div#TOC a:link {
color: blue;
}



h1.title {
  font-size: 24px;
  color: Darkblue;
  text-align: center;
  font-family: Arial, Helvetica, sans-serif;
  font-variant-caps: normal;
}
h4.author { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 {
  font-size: 22px;
  font-family: "Times New Roman", Times, serif;
  color: darkred;
  text-align: center;
}
h2 { /
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: navy;
  text-align: left;
}

h3 { 
  font-size: 15px;
  font-family: "Times New Roman", Times, serif;
  color: navy;
  text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: darkred;
  text-align: left;
}

/* unvisited link */
  a:link {
    color: green;
  }

/* visited link */
  a:visited {
    color: green;
  }

/* mouse over link */
  a:hover {
    color: red;
  }

/* selected link */
  a:active {
    color: yellow;
  }

</style>
  
  
  ```{r setup, include=FALSE}
# code chunk specifies whether the R code, warnings, and output 
# will be included in the output files.
options(repos = list(CRAN="http://cran.rstudio.com/"))

if (!require("tidyverse")) {
  install.packages("tidyverse")
  library(tidyverse)
}
if (!require("knitr")) {
  install.packages("knitr")
  library(knitr)
}
if (!require("cowplot")) {
  install.packages("cowplot")
  library(cowplot)
}
if (!require("latex2exp")) {
  install.packages("latex2exp")
  library(latex2exp)
}
if (!require("plotly")) {
  install.packages("plotly")
  library(plotly)
}
if (!require("gapminder")) {
  install.packages("gapminder")
  library(gapminder)
}
if (!require("png")) {
  install.packages("png")             # Install png package
  library("png")
}
if (!require("RCurl")) {
  install.packages("RCurl")             # Install RCurl package
  library("RCurl")
}
if (!require("colourpicker")) {
  install.packages("colourpicker")              
  library("colourpicker")
}
if (!require("gifski")) {
  install.packages("gifski")              
  library("gifski")
}
if (!require("magick")) {
  install.packages("magick")              
  library("magick")
}
if (!require("grDevices")) {
  install.packages("grDevices")              
  library("grDevices")
}
### ggplot and extensions
if (!require("ggplot2")) {
  install.packages("ggplot2")              
  library("ggplot2")
}
if (!require("gganimate")) {
  install.packages("gganimate")              
  library("gganimate")
}
if (!require("ggridges")) {
  install.packages("ggridges")              
  library("ggridges")
}
if (!require("dplyr")) {
  install.packages("dplyr")              
  library("dplyr")
}
if (!require("htmltools")) {
  install.packages("htmltools")              
  library("htmltools")
}
if (!require("rmarkdown")) {
  install.packages("rmarkdown")              
  library("rmarkdown")
}

if (!require("sf")) {
  install.packages("sf")
  library("sf")
}
if (!require("transformr")) {
  install.packages("transformr")
  library("transformr")
}
# knitr::opts_knit$set(root.dir = "/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA533/week9)

knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE)
```
```{r code, fig.width=10, fig.height=6.8}
shootings <- st_read("https://raw.githubusercontent.com/kyleknox3/STA553/main/week9/shootings.geojson") %>%
  filter(!is.na(point_x) & !is.infinite(point_x) &
         !is.na(point_y) & !is.infinite(point_y))

shootings$YEAR <- as.integer(shootings$year)


colors <- c("1" = "red", "0" = "blue")

which_state <- "pennsylvania"
county_info <- map_data("county", region = which_state)
philly_info <- county_info[county_info$subregion == "philadelphia",]


base_map <- ggplot() +
  geom_polygon(data = philly_info, aes(x = long, y = lat, group = group), color = "black", fill = "white") +
  coord_sf(xlim = c(min(shootings$point_x), max(shootings$point_x)), 
           ylim = c(min(shootings$point_y), max(shootings$point_y))) +
  theme_void()

map_with_data <- base_map +
  geom_sf(data = shootings, aes(color = as.factor(fatal)), alpha = 0.5) + 
  scale_color_manual(values = colors)

animation <- map_with_data +
  transition_time(YEAR) +
  ggtitle('Year: {frame_time}') +
  shadow_mark()

anim_save("PhillyCrimeByYear.gif", animation)

```

