---
title: "Visual Analysis of NYPD Arrests Data"
output: 
  flexdashboard::flex_dashboard:
    theme: united
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

<style type="text/css">
@import url('https://fonts.googleapis.com/css2?family=Yusei+Magic&display=swap');

body {
  background-color: lightgray;
  color: darkgreen;
  text-align: left;
}

.navbar {
  position: fixed;
  display: flex;
  justify-content: center;
  width: 100%;
}

.shiny-input-container {
  color: purple;
}

.navbg {
  background-color: blue;
  color: white;
}
</style>  

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(plotly)
library(ggplot2)
library(tidyverse)
library(leaflet)  
library(sf)

nypd_arrests <- read.csv("/Users/kyleknox/Documents/M.S. Applied Statistics/SPRING 2024/STA553/final/nypd_arrests.csv")
nypd_arrests <- na.omit(nypd_arrests)
nypd_arrests$arrest_boro <- factor(nypd_arrests$arrest_boro,
                                   levels = c("S", "K", "M", "Q", "B"),
                                   labels = c("Staten Island", "Brooklyn", "Manhattan", "Queens", "Bronx"))
nypd_arrests$perp_sex <- factor(nypd_arrests$perp_sex,
                                levels = c("M", "F"),
                                labels = c("Male", "Female"))
# Create dummy variables
sex_dummies <- model.matrix(~ perp_sex - 1, data = nypd_arrests)
boro_dummies <- model.matrix(~ arrest_boro - 1, data = nypd_arrests)

# Append the dummy variables to the original dataframe
nypd_arrests <- cbind(nypd_arrests, sex_dummies, boro_dummies)

```
Column {.sidebar .navbg data-width=200}
-------------------------------------------------------------------

<br><br>

```{r}
selectInput("selected_race", "Select Race:", choices = unique(nypd_arrests$perp_race))
selectInput("selected_gender", "Select Gender:", choices = c("Male" = "Male", "Female" = "Female"), selected = "Male")
selectInput("selected_boro", "Select Borough:", choices = c("Bronx", "Staten Island", "Brooklyn", "Manhattan", "Queens"))
```

```{r}
workData <- function() {
  if (input$selected_gender == "Male") {
    filtered_data <- nypd_arrests[nypd_arrests$perp_sex == "Male", ]
  } else if (input$selected_gender == "Female") {
    filtered_data <- nypd_arrests[nypd_arrests$perp_sex == "Female", ]
  } else {
    filtered_data <- nypd_arrests
  }
  print(head(filtered_data))
  return(filtered_data)
}

performLinearRegression <- function(data, selected_boro) {
  filtered_data <- data[data$arrest_boro == selected_boro, ]
  model <- lm(arrest_boro ~ as.numeric(age_group), data = filtered_data)
  return(model)
}

summarizeRegressionStatistics <- function(model, selected_boro) {
  filtered_data <- nypd_arrests[nypd_arrests$arrest_boro == selected_boro, ]
  summary <- summary(model)
  return(summary)
}
```

Column {data-width=450 .tabset .tabset-pills} 
-------------------------------------------------------------------
### Arrest Count by Race

```{r}
output$racePlot <- renderPlot({
  filtered_data <- nypd_arrests %>%
    filter(perp_race == input$selected_race)
  ggplot(filtered_data, aes(x = arrest_boro, fill = arrest_boro)) +
    geom_bar() +
    labs(title = paste("Arrest Count in Each Borough for", input$selected_race),
         x = "Borough", y = "Count")
})

plotOutput("racePlot")
```

### Arrest Count by Gender
```{r}
output$genderPlot <- renderPlot({
  filtered_data <- nypd_arrests %>%
    filter(perp_sex == input$selected_gender)

  ggplot(filtered_data, aes(x = arrest_boro, fill = arrest_boro)) +
    geom_bar(stat = "count") +  
    labs(title = paste("Arrest Count in Each Borough for", input$selected_gender),
         x = "Borough", y = "Count")
})

plotOutput("genderPlot")
```

### Regression Analysis
```{r}
output$regressionPlot <- renderPlot({
  data <- workData()
  selected_boro <- input$selected_boro
  filtered_data <- data[data$arrest_boro == selected_boro, ]
  model <- performLinearRegression(data, selected_boro)
  plot(filtered_data$age_group, filtered_data$arrest_boro,
       main = paste("Linear Regression Plot for", selected_boro),
       xlab = "Age Group",
       ylab = "Arrest Count")
  abline(model, lwd = 2, col = "red")
})

plotOutput("regressionPlot")
```

### Statistics Summary
```{r}
output$summaryTab <- renderPrint({
  data <- workData()
  selected_boro <- input$selected_boro
  model <- performLogisticRegression(data, selected_boro)
  summary <- summarizeInferentialStatistics(model, selected_boro)
  print(summary)
})
```

### Map Visualization

```{r}
renderLeaflet({
  filtered_data <- nypd_arrests[nypd_arrests$arrest_boro == input$selected_boro, ]

  leaflet(filtered_data) %>%
    addTiles() %>%
    addMarkers(
      lng = ~longitude, 
      lat = ~latitude, 
      popup = ~paste("Arrest Date: ", arrest_date, "<br/>",
                     "Gender: ", ifelse(perp_sex == "m", "Male", "Female"), "<br/>",
                     "Race: ", perp_race),
      clusterOptions = markerClusterOptions()
    )
})

```

Column {data-width=340}
-------------------------------------------------------------------

### Arrests by Borough

```{r}
renderPlot({
  filtered_data <- nypd_arrests[nypd_arrests$arrest_boro == input$selected_boro, ]
  ggplot(filtered_data, aes(x = perp_sex, fill = perp_race)) +
    geom_bar(position = "dodge") +
    labs(title = paste("Distribution of Arrests by Gender and Race in", input$selected_boro),
         x = "Gender", y = "Count")
})
```
